\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}


\begin{document}

\noindent Gavin Grob
\\CS 510 Automata Theory
\\Final Exam
\\
\\1. Give a regular expression for the following language where $\Sigma = \{0,1\}$. 
\\$L_1 = \{ s | s$ starts with a 0 and has even length or stats with a 1 and has odd length$\}$ 
\\Justify that your regular expression is correct by explaining why all strings generated fit the description and all strings that fit the description are generated.
\\
\\$(0(0 \cup 1) \cup 1) (00 \cup 01 \cup 10 \cup 11)$*
\\
\\I am assuming that this language doesn't accept the empty string, because the discription states that it either starts with a 0 or 1. So the regular expression doesn't accept the empty string. The expression has 2 steps, firsts generates a string starting with a 0 followed by 1 or 0 making it even, or it generates a 1 making it odd. Second it generates 2 more characters(all possible combinations of 0 and 1) keeping the generates string either even or odd accordingly.
\\
\\
\\2.Give a context free grammar generating the following language where $s$ and $t$ are strings made up of 0?s and 1?s.
\\$L_2 \{ s\#t | s^R$ is a substring of $t \}$ where $s^R$ represents the reverse of $s$.
\\Justify your grammar is correct by explaining why all strings generated fit the description and all strings that fit the description are generated.
\\
\\$S \rightarrow 1M1F | 0M0F | \# F$
\\$M \rightarrow 1M1 | 0M0 | \# F$
\\$F \rightarrow 1F | 0F | \epsilon$
\\
\\The grammar works by $1M1F  \rightarrow 10M01F  \rightarrow 10\#F01F$ then $F$ can make any string with $0$'s and $1$'s 
\\
\\So It generates $s$ and $s^R$ at the same time, then leaves a $F$ before and after $s^R$ so $t$ is able to be any string that has a substring of $s^R$
\\note: that $s$ can also be the empty string, and the $s^R$ of the empty string is the empty string. So it is valid to have the string $\# 010$ which it can also generate.
\\
\\3. Classify the following languages and concisely justify your answers.
\\
\\(a) $L_1 = \{ w | w = 0^n1^m2^m3^n$ for $ n,m \ge 1 \}$
\\i. the language is recursive but not context free
\\$\textbf{ii. the language is context free but not regular}$
\\iii. the language is regular
\\
\\A context free for $L_1$
\\$S \rightarrow 0S3 | M$
\\$M \rightarrow 1M2 | \epsilon$
\\
\\ So I know it is at least CFG, we will use the pumping lemma to prove that it is not regular
\\
\\we have a pumping length p,
\\$w \in L_1$ $w = 0^p1^{2p}2^{2p}3^p$
\\This forces y to be $1^i$ where $1 \le |xy| \le p$ since $|xy| \le n$
\\We can then unpump so we left with the string $0^{p - i}1^{2p}2^{2p}3^p$ which is not in the language
\\
\\(b) $L_2 = \{ w |$ every prefix of $w$ has more 1?s than 0?s but not more than three more 1?s than 0?s $\}$
\\$\textbf{i. the language is recursive but not context free}$
\\ii. the language is context free but not regular
\\iii. the language is regular
\\
\\I will use the CFL pumping lemma to prove $L_2$ is not context, therefore not regular.
\\
\\we have a pumping length p = 1
\\$w \in L_2$ $w = 11$
\\
\\This then forces v or x to be 1, so then if we pump there will be more than three 1's than 0's.
\\
\\
\\(c.) $L_3 = \{ <M, w > | M$ is a Turing Machine that does not read every character of its input $w \}$
\\i. recursive
\\ii. recursively enumerable but not recursive
\\$\textbf{iii. co-recursively enumerable but not recursive}$
\\iv. none of the above
\\
\\I will prove this by first showing $\overline{L_3}$ is undecidable.
\\
\\-$TM_{accept}$ = on input (M, w)
\\-$\hspace{5mm}$1.Construct  $TM_{transform}$ by modifying M on w
\\-$\hspace{10mm}$a. copy w into the input tape (this is what is read when running)
\\-$\hspace{10mm}$b. copy all the state of M into $TM_{transform}$
\\-$\hspace{10mm}$c. add a case when M goes into the accepting state, $TM_{transform}$ reads all the input.
\\-$\hspace{5mm}$2. run $TM_{read-all}$ on $TM_{transform}$
\\-$\hspace{5mm}$3. if $TM_{read-all}$ rejects then accept, otherwise reject.
\\
\\However we know that $TM_{accept}$ is undecidable, making $TM_{read-all}$ is undecidable. But like $L_u$ and $\overline{L_3}$ can list out all of the strings in this language making them RE. This then makes $L_3$  co-recursively enumerable, because $L_3$ and $\overline{L_3}$ are not RE(which would make them recursive).
\\
\\
\\4.(b) In the following one player board game you are given an $n$ x $n$ board. On each square of the board lies either a blue chip, a red chip or nothing. You win by removing chips from the board so that each column contains only chips of a single color and each row contains at least one chip. Winning may or may not be possible depending on the initial configuration of the board.
\\Let Puzzle = $\{ < B > | B$ is a winnable starting game board $\}$
\\Prove that Puzzle is $NP$-complete by first showing it is in $NP$ and then reducing an $NP$-complete problem we have studied in class to Puzzle. I suggest you use $3$-SAT as your starting point.
\\
\\Puzzle $\in$ $NP$ because we are able to scan the the columns to remove any opposite color, in less than of equal $n$ time, then scan to make sure there is still there is still a chip in the row int less than of equal $n$ time, making it polynomial time of the input size.
\\
\\3SAT $\le_p$ Puzzle
\\
\\We map our 3SAT input of $n$ variables $x_1 ... x_n$ and m clauses $c_1 ... c_m$ we make a game board $B[m,n]$:
\\1. remove any clauses that contains $x_i$ v $\overline{x_i}$.
\\2. Label each row with a clause number, 1 through m
\\3. Label each column with a variable number, 1 through n
\\4. Generating a board game $B[m,n]$
\\
\\If $x_i$ is in clause $c_j$ then place a red chip in $B[j,i]$, on the contrary if $\overline{x_i}$ is in clause $c_j$ then place a blue chip in $B[j,i]$. 
\\If $x_i$ is true, then remove the blue chips from the column $i$, on the contrary if $x_i$ is false, then remove the red chips from the column $i$. 
\\Each row with one color of chips represents a satisfied clause. 
\\Any solution to the Game can be used to find if the Puzzle B is winnable, finding the color in each column and setting the corresponding variable to true if the chip is blue and false the chip is red.
\\If the 3Sat instance is satisfiable then this game is winnable. 
\\
%\section{}
%\subsection{}



\end{document}  