\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}


\begin{document}

\noindent Gavin Grob
\\CS 510 Automata Theory
\\Homework 7
\\
\\1. Let's define a $\it useless$ $\it state$ in a Turing Machine as a state that is never entered on any input string. Consider the problem of determining whether a Turing Machine has any useless states. Formulate this problem as a language and show that it is undecidable by using a reduction.
\\
\\$TM_{useless}$ = $\{<M> | M$ is a TM $\}$ 
\\
\\$TM_{accept}$ = $\{<M, w> | M$ is a TM and w is a string. $\}$ 
\\
\\We will assume $TM_{useless}$ exists, and use it to create $TM_{accept}$. 
\\We run $TM_{accept}$ on input M, we then create $TM_{transform}$ which hard codes M and w so that w is on the input tape and $TM_{transform}$ states are equal to M states. If M accepts then $TM_{transform}$ manually  goes threw all of its states. If M does not accept(either spins forever or not in the language) then $TM_{transform}$ doesn't go into an accepting state so it has useless states. 
\\So when $TM_{transform}$ is feed to $TM_{useless}$ it will accept when M did not accept, so reject. $TM_{useless}$ will reject when M did accept, so accept.
\\
\\-$TM_{accept}$ = on input (M, w)
\\-$\hspace{5mm}$1.Construct  $TM_{transform}$ by modifying M on w
\\-$\hspace{10mm}$a. copy w into the input tape (this is what is read when running)
\\-$\hspace{10mm}$b. copy all the state of M into $TM_{transform}$
\\-$\hspace{10mm}$c. add a case when M goes into the accepting state, $TM_{transform}$ enters every state.
\\-$\hspace{5mm}$2. run $TM_{useless}$ on $TM_{transform}$
\\-$\hspace{5mm}$3. if $TM_{useless}$ rejects then accept, otherwise reject.
\\
\\However we know that $TM_{accept}$ is undecidable, so then $TM_{useless}$ cannot exist, making it undecidable.
\\
\\2. Consider the problem of determining whether a two-tape TM $\it M$ ever writes a blank symbol over a non-blank symbol on its second tape during the course of its computation on any input string. Formulate this problem as a language and show that it is undecidable by using a reduction.
\\
\\$TM_{ttbs}$ = $\{<M> | M$ is a TM$\}$
\\
\\$TM_{accept}$ = $\{<M, w> | M$ is a TM and w is a string$\}$ 
\\
\\We will assume $TM_{ttbs}$ exists, and use it to create a  $TM_{accept}$. We run $TM_{accept}$ on input M, w we then feed M,w to a $TM_{transform}$.
\\We then create $TM_{transform}$ which hard codes M and w so that w is on the input tape and $TM_{transform}$ states are equal to M states, besides when M was going to write a blank on tape 2, instead it writes a character not in the language of M, ill call it $\hat{o}$. Whenever $TM_{transform}$ reads the character $\hat{o}$ off the tape, it treats it as a blank symbol. Then when it goes into a accepting state it writes a non-blank symbol on the second tape, then write a blank over that same space (incase tape 2 is empty).
\\So when $TM_{transform}$, has a non-blank written over with a blank in the second tape, w was accepted by M. 
\\So when $TM_{transform}$ is feed to $TM_{ttbs}$ it will accept when M accepted, so accept. $TM_{ttbs}$ will reject when M rejected, so reject.
\\
\\-$TM_{accept}$ = on input (M, w)
\\-$\hspace{5mm}$1. Construct  $TM_{transform}$ by modifying M on w
\\-$\hspace{10mm}$a. copy w into the input tape (this is what is read when running)
\\-$\hspace{10mm}$b. copy all the state of M into $TM_{transform}$
\\-$\hspace{10mm}$c. when M was going to write a blank on tape 2, $TM_{transform}$ writes a $\hat{o}$  instead
\\-$\hspace{10mm}$d. when $\hat{o}$ is read off the tape, $TM_{transform}$ treats it as a blank symbol
\\-$\hspace{10mm}$e. when $TM_{transform}$ goes into a accepting state, $TM_{transform}$ writes a non-blank
\\-$\hspace{14mm}$symbol on the second tape, then write a blank over that same space
\\-$\hspace{5mm}$2. run $TM_{ttbs}$ on $TM_{transform}$
\\-$\hspace{5mm}$3. if $TM_{ttbs}$ accept then accept, otherwise reject.
\\
\\However we know that $TM_{accept}$ is undecidable, so then $TM_{useless}$ cannot exist, making it undecidable.
\\
\\
%\section{}
%\subsection{}



\end{document}  